---
title: Preliminary Plot Options 
author: AG
date: updated 051023
output: pdf_document
---

## Objective

The goal is to take the output of Export_Data.rmd to create several visualization options and draw preliminary conclusions.

## Import Data

```{r,  warning = FALSE, message = FALSE, error = FALSE}
library(tidyverse)
Ratio_Final <- as_tibble(read.csv("Exported_Data/Ratio_Final.csv"))
Frequency_RFP_Pool <- as_tibble(read.csv("Exported_Data/Frequency_RFP_Pool.csv"))
Frequency_RFP_N <- as_tibble(read.csv("Exported_Data/Frequency_RFP_N.csv"))
Frequency_GFP_Pool <- as_tibble(read.csv("Exported_Data/Frequency_GFP_Pool.csv"))
Frequency_GFP_N <- as_tibble(read.csv("Exported_Data/Frequency_GFP_N.csv"))
```

## Plots

## Validation Plot

You will want to change the siRNA in the filter to whatever wells you want to see to quickly assess if your experiment worked.

```{r,  warning = FALSE, message = FALSE, error = FALSE}
library(ggridges)

check <- Ratio_Final %>%
     filter(grepl('CTRL|LMNA|BANF1|LMNB1|SUN1', siRNA))

check$siRNA <- factor(check$siRNA, levels = c("LMNB1", "BANF1", "CTRL", "LMNA", "SUN1"))

ggplot() +
     geom_density_ridges(data = check, 
          aes(y = siRNA,
          x = Ratio_RFP,
          color = siRNA,
          fill = siRNA),
          alpha = 0.1,
          jittered_points = TRUE, 
          point_alpha = 1,
          point_shape = 21,
          point_size = 0.2) +
     scale_fill_brewer(palette = "Set1") +
     scale_color_brewer(palette = "Pastel1") +
     theme_classic() + 
     xlim(0,3) +
     xlab('nuc:cyto RFP') + 
     ylab('siRNA') +
     theme(legend.position = "none") 
     #ggsave("filepath/validation.png")
```

## Barplot

```{r,  warning = FALSE, message = FALSE, error = FALSE}
R_data_pool <- Frequency_RFP_Pool %>%
     mutate('fill' = case_when(
          siRNA == 'CTRL' ~ 'a',
          siRNA == 'SUN1' ~ 'a',
          siRNA == 'LMNB1' ~ 'a',
          siRNA == 'LMNA' ~ 'a',
          siRNA == 'BANF1' ~ 'a'
     ))

CTRL <- R_data_pool %>% filter(siRNA == 'CTRL') %>% select(rupture_frequency)
R_barplot <- R_data_pool %>%
     mutate('fold_change' = (rupture_frequency/CTRL$rupture_frequency))
R_barplot_N <- Frequency_RFP_N %>%
     mutate('fold_change' = (rupture_frequency/CTRL$rupture_frequency))

ggplot() +
     geom_col(data = R_barplot, 
          aes(x = reorder(siRNA, rupture_frequency), 
          y = rupture_frequency,
          fill = fill),
          alpha = 0.5) +
     scale_fill_manual(values = c('#000000','#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('siRNA') + 
     ylab('rupture_frequency') +
     geom_point(data = R_barplot_N,
          aes(x = siRNA,
          y = rupture_frequency,
          color = Experiment,
          size = total),
          alpha = 0.4) +
     scale_color_manual(values = c('#102343', '#78BA39', '#931491', '#523323', '#000000')) +
     theme(
          #legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
     scale_y_continuous(expand = c(0,0))
#ggsave("filepath/RFP_barplot.png")

```

### Waterfall Plot

If you want to highlight different data, you can redirect the mutate in the first chunk of code. After the graph is a commented out ggsave that you can adjust if you would like to export this plot individually.

__Loss of compartmentalization by RFP__


```{r,  warning = FALSE, message = FALSE, error = FALSE}
R_data_pool <- Frequency_RFP_Pool %>%
     mutate('fill' = case_when(
          siRNA == 'CTRL' ~ 'a',
          siRNA == 'SUN1' ~ 'a',
          siRNA == 'LMNB1' ~ 'a',
          siRNA == 'LMNA' ~ 'a',
          siRNA == 'BANF1' ~ 'a'
     ))

CTRL <- R_data_pool %>% filter(siRNA == 'CTRL') %>% select(rupture_frequency)
R_waterfall <- R_data_pool %>%
     mutate('fold_change' = (rupture_frequency/CTRL$rupture_frequency))
R_waterfall_N <- Frequency_RFP_N %>%
     mutate('fold_change' = (rupture_frequency/CTRL$rupture_frequency))

ggplot() +
     geom_col(data = R_waterfall, 
          aes(x = reorder(siRNA, log2(fold_change)), 
          y = log2(fold_change),
          fill = fill),
          alpha = 0.8) +
     scale_fill_manual(values = c('#000000','#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('siRNA') + 
     ylab('log2(fold change) nuc:cyto RFP') +
     geom_point(data = R_waterfall_N,
          aes(x = siRNA,
          y = log2(fold_change),
          color = Experiment,
          size = total),
          alpha = 0.3) +
     scale_color_manual(values = c('#102343', '#78BA39', '#931491', '#523323', '#000000')) +
     theme(
          legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
     scale_y_continuous(expand = c(0,0))
#ggsave("filepath/RFP_waterfall.png")
```

__Loss of compartmentalization by GFP normalized to rupture frequency__


```{r,  warning = FALSE, message = FALSE, error = FALSE}
G_data_pool <- Frequency_GFP_Pool %>%
     mutate('fill' = case_when(
          siRNA == 'CTRL' ~ 'a',
          siRNA == 'SUN1' ~ 'a',
          siRNA == 'LMNB1' ~ 'a',
          siRNA == 'LMNA' ~ 'a',
          siRNA == 'BANF1' ~ 'a'
     ))

CTRL <- G_data_pool %>% filter(siRNA == 'CTRL') %>% select(rupturing_frequency)
G_waterfall <- G_data_pool %>%
     mutate('fold_change' = (rupturing_frequency/CTRL$rupturing_frequency))
G_waterfall_N <- Frequency_GFP_N %>%
     mutate('fold_change' = (rupturing_frequency/CTRL$rupturing_frequency))

ggplot() +
     geom_col(data = G_waterfall, 
          aes(x = reorder(siRNA, fold_change), 
          y = log2(fold_change),
          fill = fill, 
          color = fill),
          alpha = 0.8) +
     scale_fill_manual(values = c('#000000','#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('siRNA') + 
     ylab('log2(fold change) cyto:nuc GFP')  +
     geom_point(data = G_waterfall_N,
          aes(x = siRNA,
          y = log2(fold_change),
          color = Experiment,
          size = total),
          alpha = 0.3) +
     scale_color_manual(values = c('#102343', '#78BA39', '#931491', '#523323', '#000000')) +
     theme(
          legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
     scale_y_continuous(expand = c(0,0))
     #ggsave("filepath/GFP_waterfall.png")

```
__Loss of compartmentalization by GFP normalized to rupture frequency (filtered for total rupturing cells >= 175)__


```{r,  warning = FALSE, message = FALSE, error = FALSE}
G_data_pool <- Frequency_GFP_Pool %>%
     mutate('fill' = case_when(
          siRNA == 'CTRL' ~ 'a',
          siRNA == 'SUN1' ~ 'a',
          siRNA == 'LMNB1' ~ 'a',
          siRNA == 'LMNA' ~ 'a',
          siRNA == 'BANF1' ~ 'a'
     ))

CTRL <- G_data_pool %>% filter(siRNA == 'CTRL') %>% select(rupturing_frequency)
G_waterfall <- G_data_pool %>%
     mutate('fold_change' = (rupturing_frequency/CTRL$rupturing_frequency)) %>%
     filter(total >= 175) %>%
     mutate("significant" = case_when(
          p <= 0.05/94 ~ "s",
          p > 0.05/94 ~ "ns"
     ))

G_waterfall %>%
     reframe(siRNA, total, "GFP positive" = rupturing, "p-value" = p) %>%
     kbl() %>%
     kable_paper("hover", full_width = F)

ggplot() +
     geom_col(data = G_waterfall, 
          aes(x = reorder(siRNA, fold_change), 
          y = log2(fold_change),
          fill = significant, 
          color = fill),
          alpha = 0.8) +
     scale_fill_manual(values = c('#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('siRNA') + 
     ylab('log2(fold change) cyto:nuc GFP')  +
     scale_color_manual(values = c('#102343', '#78BA39', '#931491', '#523323', '#000000')) +
     theme(
          legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
     scale_y_continuous(expand = c(0,0))
     #ggsave("filepath/GFP_total_over_175_color.png")

```
### Volcano Plot

If you want to highlight different data, you can redirect the mutate in the first chunk of code. After the graph is a commented out ggsave that you can adjust if you would like to export this plot individually.

__Loss of compartmentalization by RFP__


```{r,  warning = FALSE, message = FALSE, error = FALSE}
library(ggrepel)

labels <- R_waterfall %>%
     filter(grepl('CTRL|LMNA|LMNB1|BANF1|SUN1|CTDNEP1', siRNA))

ggplot() +
     geom_point(data = R_waterfall, 
          aes(x = log2(fold_change), 
          y = -log10(p), 
          color = fill),
          alpha = 0.8) +
     #ylim(0,18) +
     scale_color_manual(values = c('#000000','#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('log2(fold change) nuc:cyto RFP') + 
     ylab('-log10(p-value)') +
     theme(
          legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank()) +
     scale_y_continuous(expand = c(0,0)) +
       geom_label_repel(data = labels,  
                   aes(x = log2(fold_change), 
                    y = -log10(p),
                    label = siRNA),
                    nudge_y = 2,
                   size = 2)
#ggsave("filepath/RFP_volcano.png")

```

__Loss of compartmentalization by GFP__


```{r,  warning = FALSE, message = FALSE, error = FALSE}
library(ggrepel)

labels <- G_waterfall %>%
     filter(grepl('CTRL|ATP9A|LMNB1|HDAC3|NDC1|CTDNEP1|BANF1', siRNA))

ggplot() +
     geom_point(data = G_waterfall, 
          aes(x = log2(fold_change), 
          y = -log10(p), 
          color = fill),
          alpha = 0.8) +
     #xlim(-1.5,1.5) +
     scale_color_manual(values = c('#000000','#C0BDBE', '#9e0142', '#d53e4f', '#f46d43', '#fdae61', '#fee08b', '#e6f598', '#abdda4', '#3288bd', '#5e4fa2', '#241e8b', '#fcd107')) +
     theme_classic() +
     xlab('log2(fold change) cyto:nuc GFP') + 
     ylab('-log10(p-value)') +
     theme(
          legend.position = "none", 
          axis.text.x = 
               element_text(angle = 90, hjust = 1, size = 5),
          axis.ticks = element_blank()) +
     scale_y_continuous(expand = c(0,0)) +
     #ylim(0,4) +
       geom_label_repel(data = labels,  
                   aes(x = log2(fold_change), 
                    y = -log10(p),
                    label = siRNA),
                   nudge_y = 0.5,
                   size = 2)
     #ggsave("filepath/GFP_volcano.png")

```

__RFP volcano, labeled by significance (FDR)__
```{r,  warning = FALSE, message = FALSE, error = FALSE}
stats_volcano <- left_join(read.csv("data/statistics.csv"), R_waterfall) 
stats_FDR <- stats_volcano %>% filter(FDR_adjusted_p <= 0.05)
stats_label_FDR <- stats_volcano %>%
     filter(grepl("RASA1|NRM|ARID4A|NDC1|LNPK|CTDNEP1|HDAC3|MYCBP2", siRNA))

ggplot() +
     geom_point(data = stats_volcano, 
          aes(x = log2(fold_change), 
          y = -log10(FDR_adjusted_p)),
          alpha = 0.6) +
     #ylim(0,18) +
     geom_point(data = stats_FDR,
               aes(x = log2(fold_change), 
          y = -log10(FDR_adjusted_p)), 
          color = "red") +
     theme_classic() +
     xlab('log2(fold change) nuc:cyto RFP') + 
     ylab('-log10(p-value)') +
     theme(
          legend.position = "none") +
     scale_y_continuous(expand = c(0,0)) +
       geom_label_repel(data = stats_label_FDR,  
                   aes(x = log2(fold_change), 
                    y = -log10(FDR_adjusted_p),
                    label = siRNA),
                    nudge_y = 2,
                   size = 2)
#ggsave("filepath/RFP_volcano_FDRstats_label.png")

```

__RFP volcano, labeled by significance (Bonferroni)__
```{r,  warning = FALSE, message = FALSE, error = FALSE}
stats_bon <- stats_volcano %>% filter(bon_adjusted_p <= 0.05)
stats_label_bon <- stats_volcano %>%
     filter(grepl("RASA1|NRM|ARID4A|NDC1|LNPK|CTDNEP1|HDAC3|MYCBP2", siRNA))

ggplot() +
     geom_point(data = stats_volcano, 
          aes(x = log2(fold_change), 
          y = -log10(bon_adjusted_p)),
          alpha = 0.6) +
     #ylim(0,18) +
     geom_point(data = stats_bon,
               aes(x = log2(fold_change), 
          y = -log10(bon_adjusted_p)), 
          color = "red") +
     theme_classic() +
     xlab('log2(fold change) nuc:cyto RFP') + 
     ylab('-log10(p-value)') +
     theme(
          legend.position = "none") +
     scale_y_continuous(expand = c(0,0)) +
       geom_label_repel(data = stats_label_bon,  
                   aes(x = log2(fold_change), 
                    y = -log10(bon_adjusted_p),
                    label = siRNA),
                    nudge_y = 2,
                   size = 2)
#ggsave("filepath/RFP_volcano_bonstats_label.png")

ggplot() +
     geom_point(data = stats_volcano, 
          aes(x = log2(fold_change), 
          y = -log10(p)),
          alpha = 0.6) +
     #ylim(0,18) +
     geom_point(data = stats_bon,
               aes(x = log2(fold_change), 
          y = -log10(p)), 
          color = "red") +
     theme_classic() +
     xlab('log2(fold change) nuc:cyto RFP') + 
     ylab('-log10(p-value)') +
     theme(
          legend.position = "none") +
     scale_y_continuous(expand = c(0,0)) +
       geom_label_repel(data = stats_label_bon,  
                   aes(x = log2(fold_change), 
                    y = -log10(p),
                    label = siRNA),
                    nudge_y = 2,
                   size = 2)
#ggsave("filepath/RFP_volcano_bonstats_unadjustedpscale.png")

```
